---
title: "Project-Applied-LM"
author: "Nahil Sobh"
date: "October 24, 2015"
output: html_document
---

#Background

In this project, we used the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants where they were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 
More information is available at http://groupware.les.inf.puc-rio.br/har


#Goal
The goal of this  project is to predict the manner in which the participants do the exercise. 
The "classe" variable in the training set is used to predict the correctness of the exercise. 


##Loading the taining and testing Data 
```{r}
pml_training=read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
pml_test    =read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
```

##Model
We will initially use the data related to the "pitch" to build modFit1 then 
"pitch" and roll" for modFit2 of the four sensors to predict the "classe" quality variable:
(if processing time is not large we may consider the "yaw" in addition to the previous two in modFit3)

Let us find the columns corresponding to pitch, roll and yaw training:
```{r}
pitch=grep("^pitch",names(pml_training))
roll=grep("^roll",names(pml_training))
yaw=grep("^yaw",names(pml_training))
```

Now let us confirm these are indeed the columns corresponding to "pitch", "roll" and "yaw" of the four sensors in the training set:
```{r}
names(pml_training)[pitch]
names(pml_training)[roll]
names(pml_training)[yaw]
```

Here is the first model where we only used the pitch data:
```{r}
library(caret)
modFit1 = train(classe~pitch_belt+pitch_arm+pitch_dumbbell+pitch_forearm,method="rf",data=pml_training)
```

In sample error: The error rate due to predicting the "classe"" using the training set.
Out of sample error: The error rate due to predicting the "classe"" using the test set.

List below are the OOB (Out-Of-Bag) error. This correponds to out-of-sample error on the training set. 
The literature also suggest that this error is a good estimate for out-of-sample error on the test set.
```{r}
modFit1$finalModel
```

Predicted values of classe based on the pml_train data.
```{r}
predict(modFit1,pml_test)
```


Now we increase the number of predictors to reduce the oob. We add the "roll" variables to the list as shown below
```{r}
library(caret)
modFit2 = train(classe~pitch_belt+pitch_arm+pitch_dumbbell+pitch_forearm+roll_belt+roll_arm+roll_dumbbell+roll_forearm,method="rf",data=pml_training)
```


The out-of-bag error is much lower than before:
```{r}
modFit2$finalModel
```

The corresponding predicted values are:
```{r}
predict(modFit2,pml_test)
```
